---
title: "Compare independent correlations"
author: "Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
---

# Dependencies
```{r}
library(ggplot2)
library(tibble)
library(beepr)
source("./functions/theme_gar.txt")
## HC4 method to compare two independent Pearson's correlations
source("./functions/hc4.txt")
## g & h multivariate distributions
source("./functions/ghmult.txt")
## bootstrap + core corr functions
source("./functions/corfun.txt")
```

## Define plot functions
```{r}
mdf <- function(pres, nseq, rhoseq){
  nr <- length(rhoseq)
  df <- tibble(power = as.vector(pres),
             n = rep(nseq, each = nr),
             rho = factor(rep(rhoseq, length(nseq))))
df
}

plot.res.fp <- function(df, 
                        ylim = c(0, 0.1),
                        ybreaks = seq(0, 1, 0.025)){
  ggplot(df, aes(x = n, y = power, colour = rho)) + theme_gar +
  scale_colour_viridis_d(end = 0.95) +
  geom_hline(yintercept = 0.05) +
  geom_line(size = 1) +
  scale_y_continuous(breaks = ybreaks) +
  coord_cartesian(ylim = ylim) +  
  guides(colour = guide_legend(override.aes = list(size = 3))) +
  labs(x = "Sample size", 
       y = "Prop. false positives")  
}

plot.res.tp <- function(df){
  ggplot(df, aes(x = n, y = power, colour = rho)) + theme_gar +
  scale_colour_viridis_d(end = 0.95) +
  geom_line(size = 1) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  coord_cartesian(ylim = c(0, 1)) +  
  guides(colour = guide_legend(override.aes = list(size = 3))) +
  labs(x = "Sample size",
       y = "Prop. true positives")  
}
```

# 0 difference, vary rho

## g=0, h=0
```{r, eval = FALSE}
set.seed(21)
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nr <- length(rhoseq)
nsim <- 4000
es <- 0 # effect size
g <- 0
h <- 0
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  rho <- rhoseq[R]
  print(paste("sim rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(S in 1:nsim){
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      res <- twocor.all(out1[1:nseq[N],1],out1[1:nseq[N],2],out2[1:nseq[N],1],out2[1:nseq[N],2])
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff00_g0h0_varyrho.RData")
```

### Results

#### Spearman + bootstrap
```{r}
load("./data/diff00_g0h0_varyrho.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.2))
pA <- p + ggtitle("Spearman + bootstrap")
pA
```

#### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.2))
pB <- p + ggtitle("Pearson + bootstrap")
pB
```

#### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.2))
pC <- p + ggtitle("Pearson + Fisher")
pC
# ggsave("./figures/g0h0_rho0_fish.png", width = 15, height = 10, units = "cm")
```

#### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.2))
pD <- p + ggtitle("Pearson + HC4")
pD
```

#### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/diff00_g0h0_varyrho_summary.pdf", width = 30, height = 20, units = "cm")
```

## g=0, h=0.2
```{r, eval = FALSE}
set.seed(21)
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nr <- length(rhoseq)
nsim <- 4000
es <- 0 # effect size
g <- 0
h <- 0.2
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  rho <- rhoseq[R]
  print(paste("sim rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(N in 1:length(nseq)){
    n <- nseq[N]
    for(S in 1:nsim){
      out1 <- ghmul(n, g=g, h=h, p=2, cmat=cmat1)  
      out2 <- ghmul(n, g=g, h=h, p=2, cmat=cmat2)  
      res <- twocor.all(out1[,1],out1[,2],out2[,1],out2[,2])
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff00_g0h02_varyrho.RData")
```

### Results

#### Spearman + bootstrap
```{r}
load("./data/diff00_g0h02_varyrho.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.2))
pA <- p + ggtitle("Spearman + bootstrap")
pA

# ggsave("./figures/g0h02_rho0_spea.png", width = 15, height = 10, units = "cm")
```

#### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.2))
pB <- p + ggtitle("Pearson + bootstrap")
pB

# ggsave("./figures/g0h02_rho0_boot.png", width = 15, height = 10, units = "cm")
```

#### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.2))
pC <- p + ggtitle("Pearson + Fisher")
pC
# ggsave("./figures/g0h02_rho0_fish.png", width = 15, height = 10, units = "cm")
```

#### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.2))
pD <- p + ggtitle("Pearson + HC4")
pD
```

#### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/diff00_g0h02_varyrho_summary.pdf", width = 30, height = 20, units = "cm")
```

## g=0.2, h=0
```{r, eval = FALSE}
set.seed(21)
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nr <- length(rhoseq)
nsim <- 4000
es <- 0 # effect size
g <- 0.2
h <- 0
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  rho <- rhoseq[R]
  print(paste("sim rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(N in 1:length(nseq)){
    n <- nseq[N]
    for(S in 1:nsim){
      out1 <- ghmul(n, g=g, h=h, p=2, cmat=cmat1)  
      out2 <- ghmul(n, g=g, h=h, p=2, cmat=cmat2)  
      res <- twocor.all(out1[,1],out1[,2],out2[,1],out2[,2])
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff00_g02h0_varyrho.RData")
```

### Results

#### Spearman + bootstrap
```{r}
load("./data/diff00_g02h0_varyrho.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df)
pA <- p + ggtitle("Spearman + bootstrap")
pA
# ggsave("./figures/g02h0_rho0_spea.png", width = 15, height = 10, units = "cm")
```

#### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df)
pB <- p + ggtitle("Pearson + bootstrap")
pB
# ggsave("./figures/g02h0_rho0_boot.png", width = 15, height = 10, units = "cm")
```

#### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df)
pC <- p + ggtitle("Pearson + Fisher")
pC
# ggsave("./figures/g02h0_rho0_fish.png", width = 15, height = 10, units = "cm")
```

#### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df)
pD <- p + ggtitle("Pearson + HC4")
pD
```

#### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/diff00_g02h0_varyrho_summary.pdf", width = 30, height = 20, units = "cm")
```

## g=0.2, h=0.2
```{r, eval = FALSE}
set.seed(21)
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nr <- length(rhoseq)
nsim <- 4000
es <- 0 # effect size
g <- 0.2
h <- 0.2
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  rho <- rhoseq[R]
  print(paste("sim rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(N in 1:length(nseq)){
    n <- nseq[N]
    for(S in 1:nsim){
      out1 <- ghmul(n, g=g, h=h, p=2, cmat=cmat1)  
      out2 <- ghmul(n, g=g, h=h, p=2, cmat=cmat2)  
      res <- twocor.all(out1[,1],out1[,2],out2[,1],out2[,2])
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff00_g02h02_varyrho.RData")
```

### Results

#### Spearman + bootstrap
```{r}
load("./data/diff00_g02h02_varyrho.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.3), ybreaks = seq(0, 0.3, 0.05))
pA <- p + ggtitle("Spearman + bootstrap")
pA
# ggsave("./figures/g02h02_rho0_spea.png", width = 15, height = 10, units = "cm")
```

#### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.3), ybreaks = seq(0, 0.3, 0.05))
pB <- p + ggtitle("Pearson + bootstrap")
pB
# ggsave("./figures/g02h02_rho0_boot.png", width = 15, height = 10, units = "cm")
```

#### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.3), ybreaks = seq(0, 0.3, 0.05))
pC <- p + ggtitle("Pearson + Fisher")
pC
# ggsave("./figures/g02h02_rho0_fish.png", width = 15, height = 10, units = "cm")
```

#### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.3), ybreaks = seq(0, 0.3, 0.05))
pD <- p + ggtitle("Pearson + HC4")
pD
```

#### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/diff00_g02h02_varyrho_summary.pdf", width = 30, height = 20, units = "cm")
```

## g=1, h=0
```{r, eval = FALSE}
set.seed(21)
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nmax <- max(nseq)
nr <- length(rhoseq)
nsim <- 4000
es <- 0 # effect size
g <- 1
h <- 0
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  rho <- rhoseq[R]
  print(paste("sim rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(S in 1:nsim){
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      res <- twocor.all(out1[1:nseq[N],1],out1[1:nseq[N],2],out2[1:nseq[N],1],out2[1:nseq[N],2])
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff00_g1h0_varyrho.RData")
```

### Results

#### Spearman + bootstrap
```{r}
load("./data/diff00_g1h0_varyrho.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.5), ybreaks = seq(0,0.5,0.1))
pA <- p + ggtitle("Spearman + bootstrap")
pA
# ggsave("./figures/g1h0_rho0_spea.png", width = 15, height = 10, units = "cm")
```

#### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.5), ybreaks = seq(0,0.5,0.1))
pB <- p + ggtitle("Pearson + bootstrap")
pB
# ggsave("./figures/g1h0_rho0_boot.png", width = 15, height = 10, units = "cm")
```

#### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.5), ybreaks = seq(0,0.5,0.1))
pC <- p + ggtitle("Pearson + Fisher")
pC
# ggsave("./figures/g1h0_rho0_fish.png", width = 15, height = 10, units = "cm")
```

#### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.fp(df, ylim = c(0, 0.5), ybreaks = seq(0,0.5,0.1))
pD <- p + ggtitle("Pearson + HC4")
pD
```

#### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/diff00_g1h0_varyrho_summary.pdf", width = 30, height = 20, units = "cm")
```

# 0.1 difference, vary rho1

## g=0, h=0

```{r, eval = FALSE}
set.seed(21)
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nr <- length(rhoseq)
nsim <- 4000
es <- 0.1 # effect size
g <- 0
h <- 0
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  rho <- rhoseq[R]
  print(paste("sim rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(N in 1:length(nseq)){
    n <- nseq[N]
    for(S in 1:nsim){
      out1 <- ghmul(n, g=g, h=h, p=2, cmat=cmat1)  
      out2 <- ghmul(n, g=g, h=h, p=2, cmat=cmat2)  
      res <- twocor.all(out1[,1],out1[,2],out2[,1],out2[,2])
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g0h0_varyrho1.RData")
```

### Results

#### Spearman + bootstrap
```{r}
load("./data/diff01_g0h0_varyrho1.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pA <- p + ggtitle("Spearman + bootstrap")
pA
```

#### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pB <- p + ggtitle("Pearson + bootstrap")
pB
```

#### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pC <- p + ggtitle("Pearson + Fisher")
pC
# ggsave("./figures/diff01_g0h0_varyrho_fish.png", width = 15, height = 10, units = "cm")
```

#### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pD <- p + ggtitle("Pearson + HC4")
pD
```

#### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/diff01_g0h0_varyrho1_summary.pdf", width = 30, height = 20, units = "cm")
```

## g=0, h=0.2
```{r}
set.seed(21)
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nr <- length(rhoseq)
nsim <- 4000
es <- 0.1 # effect size
g <- 0
h <- 0.2
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  rho <- rhoseq[R]
  print(paste("sim rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(N in 1:length(nseq)){
    n <- nseq[N]
    for(S in 1:nsim){
      out1 <- ghmul(n, g=g, h=h, p=2, cmat=cmat1)  
      out2 <- ghmul(n, g=g, h=h, p=2, cmat=cmat2)  
      res <- twocor.all(out1[,1],out1[,2],out2[,1],out2[,2])
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g0h02_varyrho1.RData")
```

### Results

#### Spearman + bootstrap
```{r, eval = FALSE}
load("./data/diff01_g0h02_varyrho1.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pA <- p + ggtitle("Spearman + bootstrap")
pA
```

#### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pB <- p + ggtitle("Pearson + bootstrap")
pB
```

#### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pC <- p + ggtitle("Pearson + Fisher")
pC
# ggsave("./figures/diff01_g0h02_varyrho_fish.png", width = 15, height = 10, units = "cm")
```

#### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pD <- p + ggtitle("Pearson + HC4")
pD
```

#### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/diff01_g0h02_varyrho1_summary.pdf", width = 30, height = 20, units = "cm")
```

## g=0.2, h=0
```{r, eval = FALSE}
set.seed(21)
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nr <- length(rhoseq)
nsim <- 4000
es <- 0.1 # effect size
g <- 0.2
h <- 0
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  rho <- rhoseq[R]
  print(paste("sim rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(N in 1:length(nseq)){
    n <- nseq[N]
    for(S in 1:nsim){
      out1 <- ghmul(n, g=g, h=h, p=2, cmat=cmat1)  
      out2 <- ghmul(n, g=g, h=h, p=2, cmat=cmat2)  
      res <- twocor.all(out1[,1],out1[,2],out2[,1],out2[,2])
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g02h0_varyrho1.RData")
```

### Results

#### Spearman + bootstrap
```{r}
load("./data/diff01_g02h0_varyrho1.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pA <- p + ggtitle("Spearman + bootstrap")
pA
```

#### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pB <- p + ggtitle("Pearson + bootstrap")
pB
```

#### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pC <- p + ggtitle("Pearson + Fisher")
pC
```

#### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pD <- p + ggtitle("Pearson + HC4")
pD
```

#### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/diff01_g02h0_varyrho1_summary.pdf", width = 30, height = 20, units = "cm")
```

## g=0.2, h=0.2
```{r, eval = FALSE}
set.seed(21)
rhoseq <- seq(0, 0.8, 0.1)
nseq <- seq(50, 500, 50)
nr <- length(rhoseq)
nsim <- 4000
es <- 0.1 # effect size
g <- 0.2
h <- 0.2
res.spear <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, nr, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, nr, length(nseq)))

for(R in 1:nr){
  beep(2)
  rho <- rhoseq[R]
  print(paste("sim rho =",rho))
  cmat1 <- matrix(c(1,rho,rho,1),2,2)
  cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)
  for(N in 1:length(nseq)){
    n <- nseq[N]
    for(S in 1:nsim){
      out1 <- ghmul(n, g=g, h=h, p=2, cmat=cmat1)  
      out2 <- ghmul(n, g=g, h=h, p=2, cmat=cmat2)  
      res <- twocor.all(out1[,1],out1[,2],out2[,1],out2[,2])
      res.spear[S, R, N] <- res$sig.boot.spear
      res.pears.boot[S, R, N] <- res$sig.boot.pears
      res.pears.fish[S, R, N] <- res$sig.fisher.pears
      res.pears.hc4[S, R, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     rhoseq, nseq, nsim, es, g, h,
     file = "./data/diff01_g02h02_varyrho1.RData")
```

### Results

#### Spearman + bootstrap
```{r}
load("./data/diff01_g02h02_varyrho1.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pA <- p + ggtitle("Spearman + bootstrap")
pA
```

#### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pB <- p + ggtitle("Pearson + bootstrap")
pB
```

#### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pC <- p + ggtitle("Pearson + Fisher")
pC
```

#### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, rhoseq)

p <- plot.res.tp(df)
pD <- p + ggtitle("Pearson + HC4")
pD
```

#### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/diff01_g02h02_varyrho1_summary.pdf", width = 30, height = 20, units = "cm")
```

# rho1 = rho2 = 0.3, vary g, h = 0

```{r, eval = FALSE}
set.seed(21)
gseq <- seq(0, 1, 0.1)
ng <- length(gseq)
h <- 0
nseq <- seq(50, 500, 50)
nsim <- 4000

rho <- 0.3
es <- 0 # effect size
cmat1 <- matrix(c(1,rho,rho,1),2,2)
cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)

res.spear <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, ng, length(nseq)))

for(G in 1:ng){
  beep(2)
  g <- gseq[G]
  print(paste("sim g =",g))
  for(N in 1:length(nseq)){
    n <- nseq[N]
    for(S in 1:nsim){
      out1 <- ghmul(n, g=g, h=h, p=2, cmat=cmat1)  
      out2 <- ghmul(n, g=g, h=h, p=2, cmat=cmat2)  
      res <- twocor.all(out1[,1],out1[,2],out2[,1],out2[,2])
      res.spear[S, G, N] <- res$sig.boot.spear
      res.pears.boot[S, G, N] <- res$sig.boot.pears
      res.pears.fish[S, G, N] <- res$sig.fisher.pears
      res.pears.hc4[S, G, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     gseq, nseq, nsim, es, h, rho,
     file = "./data/rho03_diff00_h0_varyg.RData")
```

## Define plot functions
```{r}
mdf <- function(pres, nseq, gseq){
  ng <- length(gseq)
df <- tibble(power = as.vector(pres),
             n = rep(nseq, each = ng),
             g = factor(rep(gseq, length(nseq))))
df
}

plot.res.fp <- function(df, 
                        ylim = c(0, 0.1),
                        ybreaks = seq(0, 1, 0.025)){
  ggplot(df, aes(x = n, y = power, colour = g)) + theme_gar +
  scale_colour_viridis_d(end = 0.95) +
  geom_hline(yintercept = 0.05) +
  geom_line(size = 1) +
  scale_y_continuous(breaks = ybreaks) +
  coord_cartesian(ylim = ylim) +  
  guides(colour = guide_legend(override.aes = list(size = 3))) +
  labs(x = "Sample size", 
       y = "Prop. false positives")  
}

plot.res.tp <- function(df){
  ggplot(df, aes(x = n, y = power, colour = g)) + theme_gar +
  scale_colour_viridis_d(end = 0.95) +
  geom_line(size = 1) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  coord_cartesian(ylim = c(0, 1)) +  
  guides(colour = guide_legend(override.aes = list(size = 3))) +
  labs(x = "Sample size",
       y = "Prop. true positives")  
}
```

## Results

### Spearman + bootstrap
```{r}
load("./data/rho03_diff00_h0_varyg.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.fp(df, ylim = c(0, 0.21))
pA <- p + ggtitle("Spearman + bootstrap")
pA
# ggsave("./figures/rho03_diff00_h0_varyg_spea.png", width = 15, height = 10, units = "cm")
```

#### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.fp(df, ylim = c(0, 0.21))
pB <- p + ggtitle("Pearson + bootstrap")
pB
# ggsave("./figures/rho03_diff00_h0_varyg_boot.png", width = 15, height = 10, units = "cm")
```

### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.fp(df, ylim = c(0, 0.21))
pC <- p + ggtitle("Pearson + Fisher")
pC
# ggsave("./figures/rho03_diff00_h0_varyg_fish.png", width = 15, height = 10, units = "cm")
```

### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.fp(df, ylim = c(0, 0.21))
pD <- p + ggtitle("Pearson + HC4")
pD
```

### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/rho03_diff00_h0_varyg.pdf", width = 30, height = 20, units = "cm")
```

# rho1 = 0.3, rho2 = 0.5, vary g, h = 0

```{r, eval = FALSE}
set.seed(21)
gseq <- seq(0, 1, 0.1)
ng <- length(gseq)
h <- 0
nseq <- seq(50, 500, 50)
nsim <- 4000

rho <- 0.3
es <- 0.2 # effect size
cmat1 <- matrix(c(1,rho,rho,1),2,2)
cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)

res.spear <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, ng, length(nseq)))

for(G in 1:ng){
  beep(2)
  g <- gseq[G]
  print(paste("sim g =",g))
  for(N in 1:length(nseq)){
    n <- nseq[N]
    for(S in 1:nsim){
      out1 <- ghmul(n, g=g, h=h, p=2, cmat=cmat1)  
      out2 <- ghmul(n, g=g, h=h, p=2, cmat=cmat2)  
      res <- twocor.all(out1[,1],out1[,2],out2[,1],out2[,2])
      res.spear[S, G, N] <- res$sig.boot.spear
      res.pears.boot[S, G, N] <- res$sig.boot.pears
      res.pears.fish[S, G, N] <- res$sig.fisher.pears
      res.pears.hc4[S, G, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     gseq, nseq, nsim, es, h, rho,
     file = "./data/rho03_diff02_h0_varyg.RData")
```

## Results

### Spearman + bootstrap
```{r}
load("./data/rho03_diff02_h0_varyg.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.tp(df)
pA <- p + ggtitle("Spearman + bootstrap")
pA
# ggsave("./figures/rho03_diff02_h0_varyg_spea.png", width = 15, height = 10, units = "cm")
```

### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.tp(df)
pB <- p + ggtitle("Pearson + bootstrap")
pB
# ggsave("./figures/rho03_diff02_h0_varyg_boot.png", width = 15, height = 10, units = "cm")
```

### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.tp(df)
pC <- p + ggtitle("Pearson + Fisher")
pC
# ggsave("./figures/rho03_diff02_h0_varyg_fish.png", width = 15, height = 10, units = "cm")
```

### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.tp(df)
pD <- p + ggtitle("Pearson + HC4")
pD
```

### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/rho03_diff02_h0_varyg.pdf", width = 30, height = 20, units = "cm")
```

# rho1 = 0, rho2 = 0.6, vary g, h = 0

```{r, eval = FALSE}
set.seed(21)
gseq <- seq(0, 1, 0.1)
ng <- length(gseq)
h <- 0
nseq <- c(seq(10, 100, 10), 150, 200)
nmax <- max(nseq)
nsim <- 4000

rho <- 0
es <- 0.6 # effect size
cmat1 <- matrix(c(1,rho,rho,1),2,2)
cmat2 <- matrix(c(1,rho+es,rho+es,1),2,2)

res.spear <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.boot <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.fish <- array(0, dim = c(nsim, ng, length(nseq)))
res.pears.hc4 <- array(0, dim = c(nsim, ng, length(nseq)))

for(G in 1:ng){
  beep(2)
  g <- gseq[G]
  print(paste("sim g =",g))
  for(S in 1:nsim){
    # generate max n sample
    out1 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat1)  
    out2 <- ghmul(nmax, g=g, h=h, p=2, cmat=cmat2) 
    for(N in 1:length(nseq)){
      # subsample from max n sample
      res <- twocor.all(out1[1:nseq[N],1],out1[1:nseq[N],2],out2[1:nseq[N],1],out2[1:nseq[N],2])
      res.spear[S, G, N] <- res$sig.boot.spear
      res.pears.boot[S, G, N] <- res$sig.boot.pears
      res.pears.fish[S, G, N] <- res$sig.fisher.pears
      res.pears.hc4[S, G, N] <- res$sig.hc4.pears
    }
  }
}

beep(8)
save(res.spear, res.pears.boot, res.pears.fish, res.pears.hc4,
     gseq, nseq, nsim, es, h, rho,
     file = "./data/rho0_diff06_h0_varyg.RData")
```

## Results

### Spearman + bootstrap
```{r}
load("./data/rho0_diff06_h0_varyg.RData")

pres <- apply(res.spear, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.tp(df) + scale_x_continuous(breaks = nseq)
pA <- p + ggtitle("Spearman + bootstrap")
pA
# ggsave("./figures/rho0_diff06_h0_varyg_spea.png", width = 15, height = 10, units = "cm")
```

For a given expected population correlation difference, a certain *g* and a desired long run power value, we can use interpolation to find out the matching sample size.

To achieve at least 80% power given an expected population difference of 0.6 and *g*=0, the minimum sample size is `r round(approx(y=nseq,x=pres[1,],xout=0.8)$y)` observations. For 90% power, the minimum sample size is `r round(approx(y=nseq,x=pres[1,],xout=0.9)$y)` observations.

Now with *g*=1, to achieve at least 80% power the minimum sample size is `r round(approx(y=nseq,x=pres[11,],xout=0.8)$y)` observations; to achieve at least 90% power the minimum sample size is `r round(approx(y=nseq,x=pres[11,],xout=0.9)$y)` observations

### Pearson + bootstrap
```{r}
pres <- apply(res.pears.boot, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.tp(df) + scale_x_continuous(breaks = nseq)
pB <- p + ggtitle("Pearson + bootstrap")
pB
# ggsave("./figures/rho0_diff06_h0_varyg_boot.png", width = 15, height = 10, units = "cm")
```

### Pearson + Fisher
```{r}
pres <- apply(res.pears.fish, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.tp(df) + scale_x_continuous(breaks = nseq)
pC <- p + ggtitle("Pearson + Fisher")
pC
# ggsave("./figures/rho0_diff06_h0_varyg_fish.png", width = 15, height = 10, units = "cm")
```

To achieve at least 80% power given an expected population difference of 0.6 and *g*=0, the minimum sample size is `r round(approx(y=nseq,x=pres[1,],xout=0.8)$y)` observations. For 90% power, the minimum sample size is `r round(approx(y=nseq,x=pres[1,],xout=0.9)$y)` observations.

Now with *g*=1, to achieve at least 80% power the minimum sample size is `r round(approx(y=nseq,x=pres[11,],xout=0.8)$y)` observations; to achieve at least 90% power the minimum sample size is `r round(approx(y=nseq,x=pres[11,],xout=0.9)$y)` observations

### Pearson + HC4
```{r}
pres <- apply(res.pears.hc4, c(2,3), mean) # power
df <- mdf(pres, nseq, gseq)

p <- plot.res.tp(df) + scale_x_continuous(breaks = nseq)
pD <- p + ggtitle("Pearson + HC4")
pD
```

### Combine panels

```{r, eval = FALSE}
cowplot::plot_grid(pA, pB, pC, pD,
                   labels = c("A", "B", "C", "D"))
ggsave("./figures/rho0_diff06_h0_varyg.pdf", width = 40, height = 20, units = "cm")
```

