---
title: "Simulate correlations using g & h distributions"
author: "Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
---

# Dependencies
```{r}
library(tibble)
library(ggplot2)
library(cowplot)
source("./functions/theme_gar.txt")
```


## Define functions
```{r}
ghmul<-function(n,g=0,h=0,p=2,cmat=diag(rep(1,p))){
#
# generate n observations from a p-variate dist
# based on the g and h dist.
#
# cmat is the correlation matrix
#
x<-rmulnorm(n,p,cmat)
for(j in 1:p){
if (g>0){
x[,j]<-(exp(g*x[,j])-1)*exp(h*x[,j]^2/2)/g
}
if(g==0)x[,j]<-x[,j]*exp(h*x[,j]^2/2)
}
x
}

rmulnorm<-function(n,p,cmat){
#
# Generate data from a multivariate normal
# n= sample size
# p= number of variables
# cmat is the covariance (or correlation) matrix
#
# Method (e.g. Browne, M. W. (1968) A comparison of factor analytic
# techniques. Psychometrika, 33, 267-334.
#  Let U'U=R be the Cholesky decomposition of R. Generate independent data
#  from some dist yielding X. Then XU has population correlation matrix R
#
y<-matrix(rnorm(n*p),ncol=p)
rval<-matsqrt(cmat)
y<-t(rval%*%t(y))
y
}

matsqrt <- function(x){
  xev1<-NA
  xe <- eigen(x)
  xe1 <- xe$values
  if(all(xe1 >= 0)) {
    xev1 <- diag(sqrt(xe1))
  }
  if(is.na(xev1[1]))stop("The matrix has negative eigenvalues")
  xval1 <- cbind(xe$vectors)
  xval1i <- solve(xval1)
  y <- xval1 %*% xev1 %*% xval1i
  y
}
```

# Examples: rho = 0.5

## Get samples
```{r}
rho <- 0.5
gseq <- seq(0, 1, 0.1)
hseq <- seq(0, 0.4, 0.1)
gn <- length(gseq)
hn <- length(hseq)
n <- 200
cmat <- matrix(c(1,rho,rho,1),2,2)
m1 <- array(0, dim = c(n, 11, 5))
m2 <- array(0, dim = c(n, 11, 5))
for(G in 1:length(gseq)){
  for(H in 1:length(hseq)){
    out <- ghmul(n,g=gseq[G],h=hseq[H],p=2,cmat=cmat)
    m1[, G, H] <- out[,1]
    m2[, G, H] <- out[,2]
  }
}
```

## Make figure
```{r}
df <- tibble(x = as.vector(m1),
             y = as.vector(m2),
             g = factor(rep(rep(gseq, each = n), hn)),
             h = factor(rep(hseq, each = n*gn)))

ggplot(df, aes(x=x, y=y)) + theme_gar +
  geom_point(size = 0.5, alpha = 0.3) +
  geom_smooth(method = "lm") +
  facet_grid(rows = vars(g), cols = vars(h)) +
  coord_cartesian(xlim = c(-5, 20), ylim = c(-5, 20)) +
  theme(panel.grid.minor = element_blank(),
        axis.title = element_blank())

ggsave("./figures/gh_rho05_examples.pdf", width = 40, height = 30, units = "cm")
```

# Examples: rho = 0

## Get samples
```{r}
rho <- 0
gseq <- seq(0, 1, 0.1)
hseq <- seq(0, 0.4, 0.1)
gn <- length(gseq)
hn <- length(hseq)
n <- 200
cmat <- matrix(c(1,rho,rho,1),2,2)
m1 <- array(0, dim = c(n, 11, 5))
m2 <- array(0, dim = c(n, 11, 5))
for(G in 1:length(gseq)){
  for(H in 1:length(hseq)){
    out <- ghmul(n,g=gseq[G],h=hseq[H],p=2,cmat=cmat)
    m1[, G, H] <- out[,1]
    m2[, G, H] <- out[,2]
  }
}
```

## Make figure
```{r}
df <- tibble(x = as.vector(m1),
             y = as.vector(m2),
             g = factor(rep(rep(gseq, each = n), hn)),
             h = factor(rep(hseq, each = n*gn)))

ggplot(df, aes(x=x, y=y)) + theme_gar +
  geom_point(size = 0.5, alpha = 0.3) +
  geom_smooth(method = "lm") +
  facet_grid(rows = vars(g), cols = vars(h)) +
  coord_cartesian(xlim = c(-5, 20), ylim = c(-5, 20)) +
  theme(panel.grid.minor = element_blank(),
        axis.title = element_blank())

ggsave("./figures/gh_rho0_examples.pdf", width = 40, height = 30, units = "cm")
```

# Examples: g = 0, h = 0

## Get samples
```{r}
g <- 0 
h <- 0
rhoseq <- seq(0, 0.95, 0.05)
rn <- length(rhoseq)
n <- 200
m1 <- array(0, dim = c(n, rn))
m2 <- array(0, dim = c(n, rn))
for(R in 1:rn){
  rho <- rhoseq[R]
  cmat <- matrix(c(1,rho,rho,1),2,2)
  out <- ghmul(n,g=g,h=h,p=2,cmat=cmat)
  m1[, R] <- out[,1]
  m2[, R] <- out[,2]
}
```

## Make figure
```{r}
df <- tibble(x = as.vector(m1),
             y = as.vector(m2),
             rho = factor(rep(rhoseq, each = n)))

ggplot(df, aes(x=x, y=y)) + theme_gar +
  geom_point(size = 0.5, alpha = 0.3) +
  geom_smooth(method = "lm") +
  facet_wrap(vars(rho)) +
  coord_cartesian(xlim = c(-5, 20), ylim = c(-5, 20)) +
  theme(panel.grid.minor = element_blank(),
        axis.title = element_blank())

ggsave("./figures/gh_g0h0_examples.pdf", width = 40, height = 30, units = "cm")
```

# Examples: g = 1, h = 0

## Get samples
```{r}
g <- 1
h <- 0
rhoseq <- seq(0, 0.95, 0.05)
rn <- length(rhoseq)
n <- 200
m1 <- array(0, dim = c(n, rn))
m2 <- array(0, dim = c(n, rn))
for(R in 1:rn){
  rho <- rhoseq[R]
  cmat <- matrix(c(1,rho,rho,1),2,2)
  out <- ghmul(n,g=g,h=h,p=2,cmat=cmat)
  m1[, R] <- out[,1]
  m2[, R] <- out[,2]
}
```

## Make figure
```{r}
df <- tibble(x = as.vector(m1),
             y = as.vector(m2),
             rho = factor(rep(rhoseq, each = n)))

ggplot(df, aes(x=x, y=y)) + theme_gar +
  geom_point(size = 0.5, alpha = 0.3) +
  geom_smooth(method = "lm") +
  facet_wrap(vars(rho)) +
  coord_cartesian(xlim = c(-5, 20), ylim = c(-5, 20)) +
  theme(panel.grid.minor = element_blank(),
        axis.title = element_blank())

ggsave("./figures/gh_g1h0_examples.pdf", width = 40, height = 30, units = "cm")
```

# Examples: g = 0, h = 0.2

## Get samples
```{r}
g <- 0 
h <- 0.2
rhoseq <- seq(0, 0.95, 0.05)
rn <- length(rhoseq)
n <- 200
m1 <- array(0, dim = c(n, rn))
m2 <- array(0, dim = c(n, rn))
for(R in 1:rn){
  rho <- rhoseq[R]
  cmat <- matrix(c(1,rho,rho,1),2,2)
  out <- ghmul(n,g=g,h=h,p=2,cmat=cmat)
  m1[, R] <- out[,1]
  m2[, R] <- out[,2]
}
```

## Make figure
```{r}
df <- tibble(x = as.vector(m1),
             y = as.vector(m2),
             rho = factor(rep(rhoseq, each = n)))

ggplot(df, aes(x=x, y=y)) + theme_gar +
  geom_point(size = 0.5, alpha = 0.3) +
  geom_smooth(method = "lm") +
  facet_wrap(vars(rho)) +
  coord_cartesian(xlim = c(-5, 20), ylim = c(-5, 20)) +
  theme(panel.grid.minor = element_blank(),
        axis.title = element_blank())

ggsave("./figures/gh_g0h02_examples.pdf", width = 40, height = 30, units = "cm")
```

# Examples: g = 1, h = 0.2

## Get samples
```{r}
g <- 1
h <- 0.2
rhoseq <- seq(0, 0.95, 0.05)
rn <- length(rhoseq)
n <- 200
m1 <- array(0, dim = c(n, rn))
m2 <- array(0, dim = c(n, rn))
for(R in 1:rn){
  rho <- rhoseq[R]
  cmat <- matrix(c(1,rho,rho,1),2,2)
  out <- ghmul(n,g=g,h=h,p=2,cmat=cmat)
  m1[, R] <- out[,1]
  m2[, R] <- out[,2]
}
```

## Make figure
```{r}
df <- tibble(x = as.vector(m1),
             y = as.vector(m2),
             rho = factor(rep(rhoseq, each = n)))

ggplot(df, aes(x=x, y=y)) + theme_gar +
  geom_point(size = 0.5, alpha = 0.3) +
  geom_smooth(method = "lm") +
  facet_wrap(vars(rho)) +
  coord_cartesian(xlim = c(-5, 20), ylim = c(-5, 20)) +
  theme(panel.grid.minor = element_blank(),
        axis.title = element_blank())

ggsave("./figures/gh_g1h02_examples.pdf", width = 40, height = 30, units = "cm")
```

# Summary figure: g/h = 0/0.2

n = 5000
g = 0 / 0.2
h = 0 / 0.2

## rho = 0

### Get samples
```{r}
set.seed(21)

n <- 5000
rho <- 0
cmat <- matrix(c(1,rho,rho,1),2,2)

out <- ghmul(n,g=0,h=0,p=2,cmat=cmat)
x1 <- out[,1]
y1 <- out[,2]
out <- ghmul(n,g=0.2,h=0,p=2,cmat=cmat)
x2 <- out[,1]
y2 <- out[,2]
out <- ghmul(n,g=0,h=0.2,p=2,cmat=cmat)
x3 <- out[,1]
y3 <- out[,2]
out <- ghmul(n,g=0.2,h=0.2,p=2,cmat=cmat)
x4 <- out[,1]
y4 <- out[,2]

df <- tibble(x = c(x1, x2, x3, x4),
             y = c(y1, y2, y3, y4),
             g = factor(rep(c("g=0","g=0.2","g=0","g=0.2"), each = n)),
             h = factor(rep(c("h=0","h=0","h=0.2","h=0.2"), each = n)))
```

### Make figure
```{r}
pA <- ggplot(df, aes(x=x, y=y)) + theme_gar +
  geom_point(size = 0.5, alpha = 0.1) +
  # geom_smooth(method = "lm") +
  facet_grid(rows = vars(h), cols = vars(g)) +
  coord_cartesian(xlim = c(-10, 25), ylim = c(-10, 25)) +
  theme(axis.title = element_blank())
  # theme(panel.grid.minor = element_blank())
pA
# ggsave("./figures/gh_rho0.png", width = 20, height = 20, units = "cm")
```

## rho = 0.5

### Get samples
```{r}
set.seed(21)

n <- 5000
rho <- 0.5
cmat <- matrix(c(1,rho,rho,1),2,2)

out <- ghmul(n,g=0,h=0,p=2,cmat=cmat)
x1 <- out[,1]
y1 <- out[,2]
out <- ghmul(n,g=0.2,h=0,p=2,cmat=cmat)
x2 <- out[,1]
y2 <- out[,2]
out <- ghmul(n,g=0,h=0.2,p=2,cmat=cmat)
x3 <- out[,1]
y3 <- out[,2]
out <- ghmul(n,g=0.2,h=0.2,p=2,cmat=cmat)
x4 <- out[,1]
y4 <- out[,2]

df <- tibble(x = c(x1, x2, x3, x4),
             y = c(y1, y2, y3, y4),
             g = factor(rep(c("g=0","g=0.2","g=0","g=0.2"), each = n)),
             h = factor(rep(c("h=0","h=0","h=0.2","h=0.2"), each = n)))
```

### Make figure
```{r}
pB <- ggplot(df, aes(x=x, y=y)) + theme_gar +
  geom_point(size = 0.5, alpha = 0.1) +
  # geom_smooth(method = "lm") +
  facet_grid(rows = vars(h), cols = vars(g)) +
  coord_cartesian(xlim = c(-10, 25), ylim = c(-10, 25)) +
  theme(axis.title = element_blank())
  # theme(panel.grid.minor = element_blank())
pB
# ggsave("./figures/gh_rho05.png", width = 20, height = 20, units = "cm")
```

## Combine panels

```{r}
cowplot::plot_grid(pA, pB,
                   labels = c("A", "B"))
ggsave("./figures/gh_summary.pdf", width = 40, height = 20, units = "cm")
```

# Summary figure: vary g, h = 0

n = 5000
g = 0.1 -> 0.9
h = 0

## rho = 0

### Get samples
```{r}
set.seed(21)
n <- 5000
rho <- 0
cmat <- matrix(c(1,rho,rho,1),2,2)
h <- 0
gseq <- seq(0, 1, 0.1)
ng <- length(gseq)
res <- array(NA, dim = c(n, 2, ng))
for(G in 1:ng){
  res[,,G] <- ghmul(n,g=gseq[G],h=h,p=2,cmat=cmat)
}

df <- tibble(x = as.vector(res[,1,]),
             y = as.vector(res[,2,]),
             g = factor(rep(paste0("g=",gseq), each = n)))
```

### Make figure
```{r}
pA <- ggplot(df, aes(x=x, y=y)) + theme_gar +
  geom_point(size = 0.5, alpha = 0.1) +
  # geom_smooth(method = "lm") +
  facet_wrap(vars(g), ncol = 3) +
  coord_cartesian(xlim = c(-5, 25), ylim = c(-5, 25)) +
  theme(axis.title = element_blank())
  # theme(panel.grid.minor = element_blank())
pA
# ggsave("./figures/gvar_h0_rho0.png", width = 20, height = 20, units = "cm")
```

## rho = 0.5

### Get samples
```{r}
set.seed(21)
n <- 5000
rho <- 0.5
cmat <- matrix(c(1,rho,rho,1),2,2)
h <- 0
gseq <- seq(0, 1, 0.1)
ng <- length(gseq)
res <- array(NA, dim = c(n, 2, ng))
for(G in 1:ng){
  res[,,G] <- ghmul(n,g=gseq[G],h=h,p=2,cmat=cmat)
}

df <- tibble(x = as.vector(res[,1,]),
             y = as.vector(res[,2,]),
             g = factor(rep(paste0("g=",gseq), each = n)))
```

### Make figure
```{r}
pB <- ggplot(df, aes(x=x, y=y)) + theme_gar +
  geom_point(size = 0.5, alpha = 0.1) +
  facet_wrap(vars(g), ncol = 3) +
  coord_cartesian(xlim = c(-5, 25), ylim = c(-5, 25)) +
  theme(axis.title = element_blank())
pB
# ggsave("./figures/gvar_h0_rho05.png", width = 20, height = 20, units = "cm")
```

## Combine panels

```{r}
cowplot::plot_grid(pA, pB,
                   labels = c("A", "B"))
ggsave("./figures/gh_summary.pdf", width = 40, height = 20, units = "cm")
```