---
title: "Bayesian shift function"
author: "Guillaume A. Rousselet"
date: 2018-04-06
output:
  github_document:
    html_preview: yes
    toc: yes
    toc_depth: 2
  # pdf_document:
  #   fig_caption: yes
  #   number_sections: no
  #   toc: yes
---

# Dependencies
```{r}
library(brms)
library(tibble)
library(HDInterval)
library(bayesplot)
library(BEST)
library(bayesboot) 
library(viridis)

# install.packages("devtools")
# devtools::install_github("GRousselet/rogme")
library(rogme)
```

# Functions
```{r}
getmode <- function(x){
  mcmcDensity = density(x)
  res = mcmcDensity$x[which.max(mcmcDensity$y)]
  res
}

hd.diff <- function(df, q=0.5){
  res <- rogme::hd(df$g1, q=q) - rogme::hd(df$g2, q=q)
  res
}
```

# Make data
```{r}
set.seed(21)
n <- 100
gp1 <- rexgaussian(n, mu = 400, sigma = 100, beta = 200)
gp2 <- rexgaussian(n, mu = 250, sigma = 100, beta = 350)

df <- tibble(GP = factor(c(rep(1, length(gp1)),rep(2, length(gp2)))),
             RT = c(gp1, gp2))
summary(gp1)
summary(gp2)
round(quantile(gp1, probs = seq(.1,.9,.1)))
round(quantile(gp2, probs = seq(.1,.9,.1)))
```

## Illustrate data
```{r}
p <- ggplot(df, aes(RT)) + theme_classic() +
  geom_density(aes(colour = GP), size = 2) +
  # geom_abline(intercept=0, slope=0, colour="black") +
  scale_color_viridis(discrete = TRUE) +
  coord_cartesian(xlim = c(0, 2500)) +
  # scale_x_continuous(breaks=nvec) + 
  # scale_y_continuous(limits=c(0,22), breaks=seq(0,22,5)) +
  theme(plot.title = element_text(size=22),
        axis.title.x = element_text(size = 18),
        axis.text.x = element_text(size = 14, colour="black"),
        axis.text.y = element_text(size = 16, colour="black"),
        axis.title.y = element_text(size = 18),
        legend.key.width = unit(1.5,"cm"),
        legend.position = c(0.5,0.85),
        legend.direction = "vertical",
        legend.text=element_text(size=16),
        legend.title=element_text(size=18),
        panel.background = element_rect(fill = 'grey', colour = 'black')) +
  labs(x = "Reaction times in ms", y = "Density") +
  # guides(colour = guide_legend(override.aes = list(size=3))) + # make thicker legend lines
  ggtitle("Two exGaussian distributions") 
p

# save figure
ggsave(filename='figure_data.pdf',width=7,height=5) #path=pathname
```

```{r}
t.test(gp1, gp2)
```

## OLS fit
```{r}
olsmod <- lm(RT ~ GP, data = df)
summary(olsmod)
```

## Shift function

```{r}
df <- rogme::mkt2(gp1, gp2)
# compute shift function
sf <- rogme::shifthd_pbci(data = df, formula = obs ~ gr, nboot = 200)[,4:8]

# plot shift function
psf <- plot_sf(sf, plot_theme = 2)

# change axis labels
psf <- psf +
  labs(x = "Group 1 RT quantiles (ms)",
       y = "Group 1 - group 2 \nquantile differences (ms)")

# add labels for deciles 1 & 9
psf <- add_sf_lab(psf, sf, y_lab_nudge = 100)
psf

# save figure
ggsave(filename='figure_sf.pdf',width=7,height=5) #path=pathname
```

# Strategy 1: Bayesian bootstrap

A great explanation of the Bayesian bootstrap is provided in these two blog posts by Rasmus Bååth:

http://www.sumsar.net/blog/2015/04/the-non-parametric-bootstrap-as-a-bayesian-model/
http://www.sumsar.net/blog/2015/07/easy-bayesian-bootstrap-in-r/

There is also a movie!
http://www.sumsar.net/blog/2016/07/bayesian-bootstrap-the-movie/

```{r}
df <- tibble(g1 = gp1, g2 = gp2)
qseq <- seq(0.1, 0.9, 0.1)
bbres <- tibble(Group1 = vector(mode = "numeric", length = 9),
                Group2 = vector(mode = "numeric", length = 9),
                difference = vector(mode = "numeric", length = 9),
                ci_lower = vector(mode = "numeric", length = 9),
                ci_upper = vector(mode = "numeric", length = 9))
for(S in 1:9){
  # Harrell-Davis quantile estimate for group 1
  bbres$Group1[S] <- rogme::hd(gp1, q=qseq[S])
  # Harrell-Davis quantile estimate for group 2
  bbres$Group2[S] <- rogme::hd(gp2, q=qseq[S])
  # difference
  bbres$difference[S] <- bbres$Group1[S] - bbres$Group2[S]
  
res <- bayesboot::bayesboot(data = df, 
                            statistic = hd.diff,
                            R = 1000,
                            R2 = 1000,
                            use.weights = FALSE,
                            q = qseq[S])
tmp <- hdi(res)
bbres$ci_lower[S] <- tmp[1]
bbres$ci_upper[S] <- tmp[2]
}
```

To check bayesboot posterior distributions:
```{r}
plot(res)
summary(res)
```

## Plot results
```{r}
# plot shift function
psf <- plot_sf(bbres, plot_theme = 2)

# change axis labels
psf <- psf +
  labs(x = "Group 1 RT quantiles (ms)",
       y = "Group 1 - group 2 \nquantile differences (ms)")

# # add labels for deciles 1 & 9
# psf <- add_sf_lab(psf, sf, y_lab_nudge = 100)
psf

# save figure
ggsave(filename='figure_bbsf.pdf',width=7,height=5) #path=pathname
```

# Strategy 2: Bayesian quantile regression

## Bayesian model using quantile regression

```{r}
df <- tibble(GP = factor(c(rep(1, length(gp1)),rep(2, length(gp2)))),
             RT = c(gp1, gp2))

fit.qr <- brm(bf(RT ~ GP, sigma ~ GP, quantile = 0.5),
                  family = asym_laplace,
                  data = df, cores=4)
```

```{r, eval=FALSE}
get_prior(bf(IQ ~ Group, sigma ~ Group, quantile = 0.5),
                  family = asym_laplace,
                  data = df)
```


```{r}
fit.qr
```

```{r, eval=FALSE}
fit.qr.25 <- brm(bf(RT ~ GP, sigma ~ GP, quantile = 0.25),
                  family = asym_laplace,
                  data = df, cores=4)
```

```{r, eval=FALSE}
fit.qr.25
```

## Check chains
Get chains
```{r}
posterior <- as.array(fit.qr)
dimnames(posterior)
```

```{r}
mcmc_dens_overlay(posterior, pars = c("b_Intercept", "b_GP2", "b_sigma_Intercept", "b_sigma_GP2"))
```

```{r}
# color_scheme_set("mix-blue-red")
color_scheme_set("viridis")
mcmc_trace(posterior, pars = c("b_Intercept", "b_GP2", "b_sigma_Intercept", "b_sigma_GP2"),
           facet_args = list(ncol = 1, strip.position = "left"))
```

```{r, eval=FALSE}
color_scheme_set("red")
mcmc_intervals(posterior, pars = c("b_Intercept", "b_GP2"))
```

```{r, eval=FALSE}
mcmc_intervals(posterior, pars = c("b_sigma_Intercept", "b_sigma_GP2"))
```

```{r, eval=FALSE}
# using BEST package
x0 <- as.data.frame(fit.qr, pars = "b_Intercept")[,1]
x1 <- as.data.frame(fit.qr, pars = "b_GP2")[,1]
BEST::plotPost(x0, showMode = TRUE)
```

```{r, eval=FALSE}
BEST::plotPost(x0+x1, showMode = TRUE)
```

```{r, eval=FALSE}
round(getmode(x0))
round(hdi(x0))
round(getmode(x0+x1))
round(hdi(x0+x1))
```

## Posterior samples: posterior predictive checks
Graphical posterior predictive checks
https://cran.r-project.org/web/packages/bayesplot/vignettes/graphical-ppcs.html
```{r}
p <- brms::pp_check(fit.qr, type = "dens_overlay", nsamples = 100)
p + theme_minimal() +
  coord_cartesian(xlim = c(0, 2500))

# brms::pp_check(fit.qr, type = "error_hist")
# brms::pp_check(fit.qr, type = "scatter_avg")
# brms::pp_check(fit.qr, type = "stat_2d")
# # brms::pp_check(fit1, type = "rootogram")
# brms::pp_check(fit.qr, type = "loo_pit")
```

## Full shift function using asym_laplace

```{r, eval=FALSE}
qseq <- seq(.1, .9, .1) # quantile sequence
CM <- 0.95 # credible mass
nQ <- length(qseq)
bqrsf <- tibble(Group1 = vector(mode = "numeric", length = 9),
                Group2 = vector(mode = "numeric", length = 9),
                difference = vector(mode = "numeric", length = 9),
                ci_lower = vector(mode = "numeric", length = 9),
                ci_upper = vector(mode = "numeric", length = 9))

for(Q in 1:nQ){
  
  fit <- brm(bf(RT ~ GP, sigma ~ GP, quantile = qseq[Q]),
    family = asym_laplace,
    data = df, cores=4)
  # extract chains
  x0 <- as.data.frame(fit, pars = "b_Intercept")[,1]
  x1 <- as.data.frame(fit, pars = "b_GP2")[,1]
  
  # get median + HDI
  bqrsf$Group1[Q] <- median(x0) # condition 1 - to get mode instead: getmode(x0)
  bqrsf$Group2[Q] <- median(x0+x1) # condition 2
  
  # we could also get HDI for each group estimation
  # hdi(x0, credMass = CM) 
  # hdi(x0+x1, credMass = CM)
  
  bqrsf$difference[Q] <- median(-x1) # difference
  tmp <- hdi(-x1, credMass = CM)
  bqrsf$ci_lower[Q] <- tmp[1]
  bqrsf$ci_upper[Q] <- tmp[2]
}
save(bqrsf, file = "bqrsf.RData")
```

## Plot results
```{r}
load("bqrsf.RData")
# plot shift function
psf <- plot_sf(bqrsf, plot_theme = 2)

# change axis labels
psf <- psf +
  labs(x = "Group 1 RT quantiles (ms)",
       y = "Group 1 - group 2 \nquantile differences (ms)")

# # add labels for deciles 1 & 9
# psf <- add_sf_lab(psf, sf, y_lab_nudge = 100)
psf

# save figure
ggsave(filename='figure_bqrsf.pdf',width=7,height=5) #path=pathname
```

# Strategy 3: Bayesian model with exGaussians

```{r, eval = FALSE}
get_prior(RT ~ GP,
          family = exgaussian,
          data = df)
```

```{r, eval = FALSE}
fit.exg <- brm(bf(RT ~ GP, 
                  sigma ~ GP,
                  beta ~ GP),
                  family = exgaussian(link = "identity", 
                                      link_sigma = "log", 
                                      link_beta = "log"),
                  data = df, 
                  warmup = 1000, 
                  iter = 2000, 
                  chains = 4,
                  control = list(adapt_delta = 0.999))
save(fit.exg, file="fit.exg.RData")
```

```{r}
load("fit.exg.RData")
fit.exg
```

```{r}
posterior <- as.array(fit.exg)
mcmc_dens_overlay(posterior, pars = c("b_Intercept", "b_sigma_Intercept", "b_beta_Intercept",
                                      "b_GP2", "b_sigma_GP2", "b_beta_GP2"))
```

```{r}
x <- as.data.frame(fit.exg, pars = "b_Intercept")[,1]
BEST::plotPost(x, showMode = TRUE)
```

```{r}
x <- as.data.frame(fit.exg, pars = "b_Intercept")[,1]
x1 <- as.data.frame(fit.exg, pars = "b_GP2")[,1]
BEST::plotPost(x+x1, showMode = TRUE)
```

```{r}
x <- as.data.frame(fit.exg, pars = "b_sigma_Intercept")[,1]
BEST::plotPost(exp(x), showMode = TRUE)
```

```{r}
x <- as.data.frame(fit.exg, pars = "b_beta_Intercept")[,1]
BEST::plotPost(exp(x), showMode = TRUE)
```

## Get matrix yrep of draws from the posterior distributions
```{r}
# yrep <- posterior_predict(fit.exg, draws = 500) 
yrep <- predict(fit.exg, nsamples = 500, summary = FALSE) 
y <- df$RT
```

### Check predictions
```{r}
# color_scheme_set("brightblue")
color_scheme_set("darkgray")
ppc_dens_overlay(y[1:100], yrep[1:200, 1:100])
```

```{r}
ppc_dens_overlay(y[101:200], yrep[1:200, 101:200])
```

## Medians per group
```{r}
ppc_stat_grouped(y, yrep, stat = "median", binwidth = 1, group = df$GP)
```

## Full shift function using posterior distributions from exGaussian fits

Get matrix yrep of draws from the posterior predictive distribution
```{r}
yrep <- predict(fit.exg, nsamples = 2000, summary = FALSE) 
# ppc_dens_overlay(y[1:100], yrep[1:200, 1:100])
yrep0 <- yrep[,(1:n)]
yrep1 <- yrep[,(n+1):(n*2)]
qseq <- seq(.1, .9, .1) # quantile sequence
CM <- 0.95 # credible mass
nQ <- length(qseq)

bexgsf <- tibble(Group1 = vector(mode = "numeric", length = 9),
                Group2 = vector(mode = "numeric", length = 9),
                difference = vector(mode = "numeric", length = 9),
                ci_lower = vector(mode = "numeric", length = 9),
                ci_upper = vector(mode = "numeric", length = 9))

for(Q in 1:nQ){
  # compute quantiles using R default formula
  # x0 <- apply(yrep0, 1, quantile, probs = qseq[Q])
  # x1 <- apply(yrep1, 1, quantile, probs = qseq[Q])
  
  # compute quantiles using Harrell-Davis estimator
  x0 <- apply(yrep0, 1, rogme::hd, q = qseq[Q])
  x1 <- apply(yrep1, 1, rogme::hd, q = qseq[Q])
  
  # get median + HDI
  bexgsf$Group1[Q] <- median(x0) # condition 1 - to get mode instead: getmode(x0)
  bexgsf$Group2[Q] <- median(x1) # condition 2
  
  # we could also get HDI for each group estimation
  # hdi(x0, credMass = CM) 
  # hdi(x0+x1, credMass = CM)
  
  bexgsf$difference[Q] <- bexgsf$Group1[Q] - bexgsf$Group2[Q] # difference
  tmp <- hdi(x0-x1, credMass = CM)
  bexgsf$ci_lower[Q] <- tmp[1]
  bexgsf$ci_upper[Q] <- tmp[2]
}

save(bexgsf, file = "bexgsf.RData")
```

```{r, eval=FALSE}
# Check quantile distributions
BEST::plotPost(x0, showMode = TRUE)
mean(x0)
median(x0)

BEST::plotPost(x1, showMode = TRUE)
```

## Plot results
```{r}
load("bqrsf.RData")
# plot shift function
psf <- plot_sf(bexgsf, plot_theme = 2)

# change axis labels
psf <- psf +
  labs(x = "Group 1 RT quantiles (ms)",
       y = "Group 1 - group 2 \nquantile differences (ms)")

# # add labels for deciles 1 & 9
# psf <- add_sf_lab(psf, sf, y_lab_nudge = 100)
psf

# save figure
ggsave(filename='figure_bexgsf.pdf',width=7,height=5) #path=pathname
```
