---
title: "Estimating onsets using cluster statistics: examples"
author: "Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
    # github_document:
    # html_preview: yes
    # toc: yes
    # toc_depth: 2
---

# Dependencies
```{r, message=FALSE}
library(ggplot2)
library(tibble)
library(cowplot)
library(beepr)
library(Rfast)
library(changepoint)
source("./code/functions.R")
source("./code/theme_gar.txt")
# Edit `one_over_f` function from `primer` package to control variance (Stevens, 2009). 
# Original function is available on [GitHub](https://github.com/HankStevens/primer).
# Copyright Hank Stevens.
source("./code/one_over_f.R")
# Load template: true onset = 160 ms, F=17, max at F=26
source("./code/erp_template.R")
# R version of Matlab code from Yeung et al. 2004
source("./code/eeg_noise.R")
# to use with eeg_noise function
meanpower <- unlist(read.table("./code/meanpower.txt"))
```

# Illustrate template
```{r}
df <- tibble(x = Xf,
             y = temp2)
ggplot(df, aes(x, y)) + theme_gar +
  geom_line() + 
  labs(x = "Time in ms", y = "Signal in arbitrary units")

ggsave(filename = "./figures/template.pdf", width = 10, height = 5) 
```

# Generate noise using `one_over_f` function

## For one gamma value 

```{r}
set.seed(666)
Nt <- 10 # number of trials
gsp <- 1 # gamma spectral power -- 0 = white noise, 1 = pink noise
outvar <- 1
resmat <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  resmat[T,] <- one_over_f(gamma = gsp, Nf, outvar = outvar)
}

df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(resmat),
             trials = factor(rep(seq(1,Nt), Nf)))

ggplot(df, aes(x, y, group = trials, colour = trials)) + theme_gar + 
  geom_line(show.legend = FALSE) +
  labs(x = "Time in ms", y = "Signal in arbitrary units") +
  scale_colour_viridis_d()
```

## For multiple gamma values
```{r}
gsp <- seq(0, 2, 0.25) # gamma spectral power
outvar <- 1 # single-trial noise variance
Nt <- length(gsp)
resmat <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  set.seed(1) # same random seed to better compare trials
  resmat[T,] <- one_over_f(gamma = gsp[T], Nf, outvar = outvar)
}

df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(resmat),
             gamma = factor(rep(gsp, Nf)))

ggplot(df, aes(x, y, group = gamma)) + theme_gar +
  geom_line(linewidth = 1) + 
  # scale_colour_viridis_d(end = 0.95)
  facet_wrap(vars(gamma)) +
  labs(x = "Time in ms", y = "Signal in arbitrary units")

ggsave(filename = "./figures/gamma_noise.pdf", width = 10, height = 5)
```

### Superimposed
```{r}
gsp <- seq(0, 2, 0.2) # gamma spectral power
outvar <- 1 # single-trial noise variance
Nt <- length(gsp)
resmat <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  set.seed(1) # same random seed to better compare trials
  resmat[T,] <- one_over_f(gamma = gsp[T], Nf, outvar = outvar)
}

df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(resmat),
             gamma = factor(rep(gsp, Nf)))
ggplot(df, aes(x, y, group = gamma, colour = gamma)) + theme_gar +
  geom_line(linewidth = 1) + 
  scale_colour_viridis_d(end = 0.95) +
  labs(x = "Time in ms", y = "Signal in arbitrary units")
```


## Illustrate signal + noise

```{r}
set.seed(666)
gsp <- 1 # gamma spectral power
outvar <- 1 # noise variance
Nt <- 50
cond <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  cond[T,] <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
}

df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(cond),
             trials = factor(rep(1:Nt, Nf))
             )

df2 <- tibble(x = Xf,
              y = apply(cond, 2, mean),
              trials = factor(rep(200, Nf))
              )

ggplot(df, aes(x, y, group = trials)) + theme_gar +
  geom_line(colour = "grey") +
  geom_hline(yintercept = 0) +
  geom_line(data = df2, aes(x, y), linewidth = 2, colour = "black") + 
  labs(x = "Time in ms", y = "Signal in arbitrary units")

  ggsave(filename = "./figures/gamma_noise_trials_mean.pdf", width = 10, height = 5) 
```

## Illustrate averages

```{r}
# set.seed(666)
Nt <- 50 # number of trials
gsp <- 1 # gamma spectral power
outvar <- 1 # noise variance
cond1 <- matrix(0, nrow = Nt, ncol = Nf)
cond2 <- matrix(0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  cond2[T,] <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
  cond1[T,] <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar)
}
cond1 <- apply(cond1, 2, mean)
cond2 <- apply(cond2, 2, mean)

df <- tibble(x = rep(Xf,2),
             y = c(cond1, cond2),
             conds = factor(rep(c("Condition 1", "Condition 2"), each = Nf)))
ggplot(df, aes(x, y, colour = conds)) + theme_gar +
  geom_hline(yintercept = 0) +
  geom_line(linewidth = 2, show.legend = FALSE) +
  geom_vline(xintercept = true_onset) +
  labs(x = "Time in ms", y = "Signal in arbitrary units") +
  scale_colour_manual(values = c("grey", "black"))
  
  ggsave(filename = "./figures/gamma_noise_mean.pdf", width = 10, height = 5)
```

# Generate noise using `eeg_noise` function

Use more realistic noise generated by mixing noise at different frequencies.

## Multiple trials

```{r}
set.seed(666)
Nt <- 10 # number of trials
outvar <- 1
resmat <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  resmat[T,] <- eeg_noise(frames = Nf, srate = 100, outvar = outvar, meanpower)
}

df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(resmat),
             trials = factor(rep(seq(1,Nt), Nf)))

ggplot(df, aes(x, y, group = trials, colour = trials)) + theme_gar + 
  geom_line(show.legend = FALSE) +
  labs(x = "Time in ms", y = "Signal in arbitrary units") +
  scale_colour_viridis_d()

ggsave(filename = "./figures/eeg_noise_trials.pdf", width = 10, height = 5)
```

## Multiple trials

```{r}
set.seed(666)
Nt <- 10 # number of trials
outvar <- 1
resmat <- matrix(data = 0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  resmat[T,] <- eeg_noise(frames = Nf, srate = 100, outvar = outvar, meanpower)
}

df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(resmat),
             trials = factor(rep(seq(1,Nt), Nf)))

ggplot(df, aes(x, y, group = trials, colour = trials)) + theme_gar + 
  geom_line(show.legend = FALSE) +
  labs(x = "Time in ms", y = "Signal in arbitrary units") +
  facet_wrap(vars(trials), nrow = 2)
```

## Illustrate signal + noise

```{r}
set.seed(666)
Nt <- 50 # number of trials
outvar <- 1 # noise variance
cond1 <- matrix(0, nrow = Nt, ncol = Nf)
cond2 <- matrix(0, nrow = Nt, ncol = Nf)
for(T in 1:Nt){
  cond2[T,] <- temp2 + eeg_noise(frames = Nf, srate = 100, outvar = outvar, meanpower)
  cond1[T,] <- temp1 + eeg_noise(frames = Nf, srate = 100, outvar = outvar, meanpower)
}
df <- tibble(x = rep(Xf, each = Nt),
             y = as.vector(cond2),
             trials = factor(rep(1:Nt, Nf))
             )

df2 <- tibble(x = Xf,
              y = apply(cond2, 2, mean),
              trials = factor(rep(200, Nf))
              )

ggplot(df, aes(x, y, group = trials)) + theme_gar +
  geom_line(colour = "grey") +
  geom_hline(yintercept = 0) +
  geom_line(data = df2, aes(x, y), linewidth = 2, colour = "black") + 
  labs(x = "Time in ms", y = "Signal in arbitrary units")

  ggsave(filename = "./figures/eeg_noise_trials_mean.pdf", width = 10, height = 5) 
```

## Illustrate averages

```{r}
df <- tibble(x = rep(Xf,2),
             y = c(apply(cond1, 2, mean), 
                   apply(cond2, 2, mean)),
             conds = factor(rep(c("Condition 1", "Condition 2"), each = Nf)))

ggplot(df, aes(x, y, colour = conds)) + theme_gar +
  geom_hline(yintercept = 0) +
  geom_line(linewidth = 2, show.legend = FALSE) +
  geom_vline(xintercept = true_onset) +
  labs(x = "Time in ms", y = "Signal in arbitrary units") +
  scale_colour_manual(values = c("grey", "black"))
  
  ggsave(filename = "./figures/eeg_noise_mean.pdf", width = 10, height = 5)
```


# T^2 time courses

```{r}
# t-tests
ori.t2 <- vector(mode = "numeric", length = Nf)
for(F in 1:Nf){
  ori.t2[F] <- t.test(cond1[,F], cond2[,F])$statistic^2
}

df <- tibble(x = Xf,
             y = t2)
ggplot(df, aes(x, y)) + theme_gar +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = true_onset) +
    labs(x = "Time in ms", y = bquote(t^2))
  
  ggsave(filename = "./figures/eeg_t2.pdf", width = 10, height = 5)
```

# Estimate onsets: examples

## `changepoint` package
```{r, warning=FALSE}
res <- cpt.meanvar(ori.t2, method = "BinSeg", Q=2)

df <- tibble(x = Xf,
             y = ori.t2)

p.cp <- ggplot(df, aes(x, y)) + theme_gar + 
  geom_line(linewidth = 1) +
  geom_vline(xintercept = true_onset) +
  geom_vline(xintercept = Xf[res@cpts[1]], linetype = "dotted") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("Change point onset =", Xf[res@cpts[1]], "ms"))
p.cp
```

## Permutation

### Generate distribution
```{r}
nboot <- 500
aath <- 0.05 # arbitrary alpha threshold
# Make permutation table of t values 
perm.tvals <- permtdist(cond1, cond2, Nt, Nf, nboot = nboot)
perm.t2 <- perm.tvals^2
```

### Compute onsets
```{r}
perm.th <- apply(perm.t2, 2, quantile, probs = 1-aath)

# FDR -----
perm.pvals <- vector(mode = "numeric", length = Nf)
for(F in 1:Nf){
  perm.pvals[F] <- (sum(perm.t2[,F] >= ori.t2[F]) + 1) / (nboot + 1)
}
# fdr.pvals <- p.adjust(ori.pvals, method = "fdr")
fdr.pvals <- p.adjust(perm.pvals, method = "fdr")
onset.fdr <- Xf[which(fdr.pvals <= aath)[1]]

# MAX -----
max.th <- quantile(apply(perm.t2, 1, max), probs = 1-aath)
onset.max <- Xf[which(ori.t2 >= max.th)[1]]

# cluster-sum statistics -----
cmap <- cluster.make(perm.pvals <= aath)
perm.max.sums <- vector(mode = "numeric", length = nboot)
for(B in 1:nboot){
  # threshold permutation t2 values and form clusters
  perm.cmap <- cluster.make(perm.t2[B,] <= perm.th)  
  perm.max.sums[B] <- max(cluster.sum(values = perm.t2[B,], cmap = perm.cmap))
}
# cluster sum threshold
cs.th <- quantile(perm.max.sums, probs = 1-aath)
# cluster test
cs.test <- cluster.test(values = perm.t2, cmap = cmap, cs.th)
onset.cs <- Xf[cs.test][1]
# onset.cs
```

### Plot permutation distribution + FDR onset

Superimpose original t^2 time course and 95th quantiles of permutation distribution.

```{r, warning=FALSE}
df <- tibble(x = Xf,
             y = ori.t2,
             boot = factor(rep(1, Nf)))

df.perm <- tibble(x = rep(Xf, each = nboot),
                  y = as.vector(perm.t2),
                  boot = factor(rep(1:nboot, Nf)))

df.q <- tibble(x = Xf, 
               y = perm.th,
               boot = factor(rep(1, Nf)))

df.sig <- tibble(x = Xf,
                 y = ifelse(fdr.pvals <= aath, 0, NA),
                 boot = factor(rep(1, Nf)))

p.fdr <- ggplot(data = df.perm, aes(x=x, y=y, group = boot)) + theme_gar +
  # plot permutation distributions
  geom_line(data = df.perm, aes(x=x, y=y, group = boot),
            colour = "grey",
            linewidth = 0.5) +
  # superimpose original t2 time course
  geom_line(data = df, linewidth = 1) +
  geom_vline(xintercept = true_onset) + 
  geom_line(data = df.q, linetype = "dashed") +
  # plot onset
  geom_vline(xintercept = onset.fdr, linetype = "dotted") +
  # plot statistical significance
  geom_point(data = df.sig, colour = "black") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("Permutation + FDR correction =",onset.fdr,"ms"))
p.fdr
```

### MAX correction
```{r, warning=FALSE}
df <- tibble(x = Xf,
             y = ori.t2,
             boot = factor(rep(1, Nf)))

df.perm <- tibble(x = rep(Xf, each = nboot),
                  y = as.vector(perm.t2),
                  boot = factor(rep(1:nboot, Nf)))

df.q <- tibble(x = Xf, 
               y = perm.th,
               boot = factor(rep(1, Nf)))

df.sig <- tibble(x = Xf,
                 y = ifelse(ori.t2 >= max.th, 0, NA),
                 boot = factor(rep(1, Nf)))

p.max <- ggplot(data = df.perm, aes(x=x, y=y, group = boot)) + theme_gar +
  # plot permutation distributions
  geom_line(data = df.perm, aes(x=x, y=y, group = boot),
            colour = "grey",
            linewidth = 0.5) +
  # superimpose original t2 time course
  geom_line(data = df, linewidth = 1) +
  geom_vline(xintercept = true_onset) + 
  geom_hline(yintercept = max.th, linetype = "dashed") +
  # plot onset
  geom_vline(xintercept = onset.max, linetype = "dotted") +
  # plot statistical significance
  geom_point(data = df.sig, colour = "black") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("Permutation + MAX correction =",onset.max,"ms"))
p.max
```

### cluster-sum statistics
```{r, warning=FALSE}
df <- tibble(x = Xf,
             y = ori.t2,
             boot = factor(rep(1, Nf)))

df.perm <- tibble(x = rep(Xf, each = nboot),
                  y = as.vector(perm.t2),
                  boot = factor(rep(1:nboot, Nf)))

df.q <- tibble(x = Xf, 
               y = perm.th,
               boot = factor(rep(1, Nf)))

df.sig <- tibble(x = Xf,
                 y = ifelse(cs.test, 0, NA),
                 boot = factor(rep(1, Nf)))

p.cs <- ggplot(data = df.perm, aes(x=x, y=y, group = boot)) + theme_gar +
  # plot permutation distributions
  geom_line(data = df.perm, aes(x=x, y=y, group = boot),
            colour = "grey",
            linewidth = 0.5) +
  # superimpose original t2 time course
  geom_line(data = df, linewidth = 1) +
  geom_vline(xintercept = true_onset) + 
  geom_line(data = df.q, linetype = "dashed") +
  # plot onset
  geom_vline(xintercept = onset.cs, linetype = "dotted") +
  # plot statistical significance
  geom_point(data = df.sig, colour = "black") +
  labs(x = "Time in ms", y = bquote(t^2)) +
  ggtitle(paste("Permutation + cluster-sum correction =",onset.cs,"ms"))
p.cs
```

### Combine panels

```{r, warning = FALSE}
cowplot::plot_grid(p.fdr, p.max, p.cs, p.cp,
                   nrow = 2,
                   labels = c("A", "B", "C", "D"),
                   label_size = 20)

ggsave(filename = "./figures/eeg_onset_summary.pdf", width = 15, height = 10)
```

# References

Stevens, M.H.H. (2009) A Primer of Ecology with R, 2nd Printing. Springer, New York.

Yeung, N., Bogacz, R., Holroyd, C.B., & Cohen, J.D. (2004) Detection of synchronized oscillations in the electroencephalogram: An evaluation of methods. Psychophysiology, 41, 822–832.






