---
title: "Estimating onsets using cluster statistics: simulation"
author: "Guillaume A. Rousselet"
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: no
    number_sections: no
    toc: yes
    toc_depth: 2
    # github_document:
    # html_preview: yes
    # toc: yes
    # toc_depth: 2
---

# Dependencies
```{r, message=FALSE}
library(ggplot2)
library(tibble)
library(changepoint)
library(cowplot)
library(beepr)
library(Rfast)
source("./code/functions.R")
source("./code/theme_gar.txt")
# Edit `one_over_f` function from `primer` package to control variance (Stevens, 2009). 
# Original function is available on [GitHub](https://github.com/HankStevens/primer).
# Copyright Hank Stevens.
source("./code/one_over_f.R")
# Load template: true onset = 160 ms, F=17, max at F=26
source("./code/erp_template.R")
```

# Simulation: one gamma value

```{r, eval=FALSE, warning=FALSE}
set.seed(666)
aath <- 0.05 # arbitrary alpha threshold
nsim <- 10000 # simulation iterations
nboot <- 2000 # number of permutation samples
inc.step <- 500 # console notification every inc.step iterations
simres.cp <- vector(mode = "numeric", length = nsim) * NA
simres.fdr <- vector(mode = "numeric", length = nsim) * NA
simres.max <- vector(mode = "numeric", length = nsim) * NA 
simres.cs <- vector(mode = "numeric", length = nsim) * NA

Nt <- 50 # number of trials
gsp <- 1 # gamma spectral power
outvar <- 1 # noise variance
cond1 <- matrix(0, nrow = Nt, ncol = Nf)
cond2 <- matrix(0, nrow = Nt, ncol = Nf)

for(S in 1:nsim){
  
  sim.counter(S, nsim, inc = inc.step)
  
  for(T in 1:Nt){
    cond2[T,] <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
    cond1[T,] <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar)  
  }
  
  # t-tests
  ori.t2 <- vector(mode = "numeric", length = Nf)
  for(F in 1:Nf){
    ori.t2[F] <- t.test(cond1[,F], cond2[,F])$statistic^2
  }
  # fit change point model
  res <- cpt.meanvar(ori.t2, method = "BinSeg", Q=2)
  simres.cp[S] <- Xf[res@cpts[1]]
  
  # Make permutation table of t values 
  perm.t2 <- permtdist(cond1, cond2, Nt, Nf, nboot = nboot)^2
  perm.th <- apply(perm.t2, 2, quantile, probs = 1-aath)
  
  # FDR -----
  perm.pvals <- vector(mode = "numeric", length = Nf)
  for(F in 1:Nf){
    perm.pvals[F] <- (sum(perm.t2[,F] >= ori.t2[F]) + 1) / (nboot + 1)
  }
  fdr.pvals <- p.adjust(perm.pvals, method = "fdr")
  simres.fdr[S] <- Xf[which(fdr.pvals <= aath)[1]]
  
  # MAX -----
  max.th <- quantile(apply(perm.t2, 1, max), probs = 1-aath)
  simres.max[S] <- Xf[which(ori.t2 >= max.th)[1]]
  
  # cluster-sum statistics -----
  cmap <- cluster.make(perm.pvals <= aath)
  perm.max.sums <- vector(mode = "numeric", length = nboot)
  for(B in 1:nboot){
    # threshold permutation t2 values and form clusters
    perm.cmap <- cluster.make(perm.t2[B,] <= perm.th)  
    perm.max.sums[B] <- max(cluster.sum(values = perm.t2[B,], cmap = perm.cmap))
  }
  # cluster sum threshold
  cs.th <- quantile(perm.max.sums, probs = 1-aath)
  # cluster test
  cs.test <- cluster.test(values = ori.t2, cmap = cmap, cs.th)
  simres.cs[S] <- Xf[cs.test][1]
}

save(simres.cs, simres.max, simres.fdr, simres.cp,
     file = "./data/onsetsim_n50_gamma1.RData")
```

## Plot onset distributions

```{r, warning=FALSE}
load("./data/onsetsim_n50_gamma1.RData")
# remove missing values
# simres.cs <- simres.cs[is.finite(simres.cs)]
# simres.cp <- simres.cp[is.finite(simres.cp)]
# simres.max <- simres.max[is.finite(simres.max)]
# simres.fdr <- simres.fdr[is.finite(simres.fdr)]

# Colour palette from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
categ.palette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

df <- tibble(onsets = c(simres.cp, simres.cs, simres.fdr, simres.max),
             method = factor(c(rep("change point", length(simres.cp)),
                               rep("cluster sum", length(simres.cs)),
                               rep("FDR", length(simres.fdr)),
                               rep("MAX", length(simres.max))))
)

ggplot(data = df, aes(x = onsets, colour = method)) + theme_gar +
  # stat_density(geom = "line") +
  geom_freqpoly(fill = "white", na.rm = TRUE, breaks = Xf) +
  geom_vline(xintercept = true_onset, linetype = "dashed") +
  # geom_vline(xintercept = median(simres.cp, na.rm = TRUE))
  scale_colour_manual(values = categ.palette) +
  theme(legend.position = c(.8, .8)) +
  labs(x = "Onsets in ms", y = "Count")
```

## Bias
```{r}
print("Bias:")
print(paste("FDR =",median(simres.fdr, na.rm = TRUE) - true_onset))
print(paste("MAX =",median(simres.max, na.rm = TRUE) - true_onset))
print(paste("Cluster =",median(simres.cs, na.rm = TRUE) - true_onset))
print(paste("Change point =",median(simres.cp, na.rm = TRUE) - true_onset))
```

## Mean absolute error 

```{r}
print("MAE:")
print(paste("FDR =",round(mean(abs(simres.fdr - true_onset), na.rm = TRUE), digits=0)))
print(paste("MAX =",round(mean(abs(simres.max - true_onset), na.rm = TRUE), digits=0)))
print(paste("Cluster =",round(mean(abs(simres.cs - true_onset), na.rm = TRUE), digits=0)))
print(paste("Change point =",round(mean(abs(simres.cp - true_onset), na.rm = TRUE), digits=0)))
```

## Proportion too early
```{r}
print("Proportion too early:")
print(paste("FDR =",round(100*mean((simres.fdr - true_onset) < 0, na.rm = TRUE), digits=1),"%"))
print(paste("MAX =",round(100*mean((simres.max - true_onset) < 0, na.rm = TRUE), digits=1),"%"))
print(paste("Cluster =",round(100*mean((simres.cs - true_onset) < 0, na.rm = TRUE), digits=1),"%"))
print(paste("Change point =",round(100*mean((simres.cp - true_onset) < 0, na.rm = TRUE), digits=1),"%"))
```

## Underestimations of at least 40 ms
```{r}
print("Underestimations of at least 40 ms:")
print(paste("FDR =",round(100*mean((simres.fdr - true_onset) <= -40, na.rm = TRUE), digits=1),"%"))
print(paste("MAX =",round(100*mean((simres.max - true_onset) <= -40, na.rm = TRUE), digits=1),"%"))
print(paste("Cluster =",round(100*mean((simres.cs - true_onset) <= -40, na.rm = TRUE), digits=1),"%"))
print(paste("Change point =",round(100*mean((simres.cp - true_onset) <= -40, na.rm = TRUE), digits=1),"%"))
```

# Simulation: vary gamma

Here we vary gamma from zero to two in steps of 0.2

```{r, eval=FALSE, warning=FALSE}

aath <- 0.05 # arbitrary alpha threshold
nsim <- 10000 # simulation iterations
nboot <- 2000 # number of permutation samples
inc.step <- 1 # console notification every inc.step iterations
gamma_vec <- seq(0,2,0.2) # gamma from 0 = white noise to 1 = pink noise
gamma_length <- length(gamma_vec)

simres.cp <- matrix(NA, nrow = gamma_length, ncol = nsim)
simres.fdr <- matrix(NA, nrow = gamma_length, ncol = nsim)
simres.max <- matrix(NA, nrow = gamma_length, ncol = nsim)
simres.cs <- matrix(NA, nrow = gamma_length, ncol = nsim)

Nt <- 50 # number of trials
outvar <- 1 # noise variance
cond1 <- matrix(0, nrow = Nt, ncol = Nf)
cond2 <- matrix(0, nrow = Nt, ncol = Nf)

for(G in 1:gamma_length){
  
  set.seed(666) # set seed here so results are directly comparable across gamma values    
  gsp <- gamma_vec[G] # gamma spectral power
  sim.counter(G, gamma_length, inc = inc.step)
  
  for(S in 1:nsim){
    
    for(T in 1:Nt){
      cond2[T,] <- temp2 + one_over_f(gamma = gsp, Nf, outvar = outvar)
      cond1[T,] <- temp1 + one_over_f(gamma = gsp, Nf, outvar = outvar)  
    }
    
    # t-tests
    ori.t2 <- vector(mode = "numeric", length = Nf)
    for(F in 1:Nf){
      ori.t2[F] <- t.test(cond1[,F], cond2[,F])$statistic^2
    }
    # fit change point model
    res <- cpt.meanvar(ori.t2, method = "BinSeg", Q=2)
    simres.cp[G,S] <- Xf[res@cpts[1]]
    
    # Make permutation table of t values 
    perm.t2 <- permtdist(cond1, cond2, Nt, Nf, nboot = nboot)^2
    perm.th <- apply(perm.t2, 2, quantile, probs = 1-aath)
    
    # FDR -----
    perm.pvals <- vector(mode = "numeric", length = Nf)
    for(F in 1:Nf){
      perm.pvals[F] <- (sum(perm.t2[,F] >= ori.t2[F]) + 1) / (nboot + 1)
    }
    fdr.pvals <- p.adjust(perm.pvals, method = "fdr")
    simres.fdr[G,S] <- Xf[which(fdr.pvals <= aath)[1]]
    
    # MAX -----
    max.th <- quantile(apply(perm.t2, 1, max), probs = 1-aath)
    simres.max[G,S] <- Xf[which(ori.t2 >= max.th)[1]]
    
    # cluster-sum statistics -----
    cmap <- cluster.make(perm.pvals <= aath)
    perm.max.sums <- vector(mode = "numeric", length = nboot)
    for(B in 1:nboot){
      # threshold permutation t2 values and form clusters
      perm.cmap <- cluster.make(perm.t2[B,] <= perm.th)  
      perm.max.sums[B] <- max(cluster.sum(values = perm.t2[B,], cmap = perm.cmap))
    }
    # cluster sum threshold
    cs.th <- quantile(perm.max.sums, probs = 1-aath)
    # cluster test
    cs.test <- cluster.test(values = ori.t2, cmap = cmap, cs.th)
    simres.cs[G,S] <- Xf[cs.test][1]
  }
}

save(simres.cs, simres.max, simres.fdr, simres.cp,
     file = "./data/onsetsim_t2_n50_var1_gammavar.RData")
```

## Results

Plot results as a function of gamma values.

### Compute summary statistics
```{r}
load(file = "./data/onsetsim_t2_n50_var1_gammavar.RData")

gamma_vec <- seq(0,2,0.2) # gamma from 0 = white noise to 1 = pink noise
gamma_length <- length(gamma_vec)

res.bias <- matrix(0, nrow = 4, ncol = gamma_length) 
res.mae <- matrix(0, nrow = 4, ncol = gamma_length)
res.pte <- matrix(0, nrow = 4, ncol = gamma_length)
res.p40 <- matrix(0, nrow = 4, ncol = gamma_length)

for(G in 1:gamma_length){
  #Bias
  res.bias[1,G] <- median(simres.fdr[G,], na.rm = TRUE) - true_onset
  res.bias[2,G] <- median(simres.max[G,], na.rm = TRUE) - true_onset
  res.bias[3,G] <- median(simres.cs[G,], na.rm = TRUE) - true_onset
  res.bias[4,G] <- median(simres.cp[G,], na.rm = TRUE) - true_onset
  
  #Mean absolute error 
  res.mae[1,G] <- mean(abs(simres.fdr[G,] - true_onset), na.rm = TRUE)
  res.mae[2,G] <- mean(abs(simres.max[G,] - true_onset), na.rm = TRUE)
  res.mae[3,G] <- mean(abs(simres.cs[G,] - true_onset), na.rm = TRUE)
  res.mae[4,G] <- mean(abs(simres.cp[G,] - true_onset), na.rm = TRUE)
  
  #Proportion too early
  res.pte[1,G] <- mean((simres.fdr[G,] - true_onset) < 0, na.rm = TRUE)
  res.pte[2,G] <- mean((simres.max[G,] - true_onset) < 0, na.rm = TRUE)
  res.pte[3,G] <- mean((simres.cs[G,] - true_onset) < 0, na.rm = TRUE)
  res.pte[4,G] <- mean((simres.cp[G,] - true_onset) < 0, na.rm = TRUE)
  
  #Underestimations of at least 40 ms
  res.p40[1,G] <- mean((simres.fdr[G,] - true_onset) <= -40, na.rm = TRUE)
  res.p40[2,G] <- mean((simres.max[G,] - true_onset) <= -40, na.rm = TRUE)
  res.p40[3,G] <- mean((simres.cs[G,] - true_onset) <= -40, na.rm = TRUE)
  res.p40[4,G] <- mean((simres.cp[G,] - true_onset) <= -40, na.rm = TRUE)
}
```

### Make figures

#### Bias

Cluster sum = FDR = 30 ms.
```{r}
# Colour palette from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
categ.palette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

df <- tibble(res = as.vector(res.bias),
             gamma = rep(gamma_vec, each = 4),
             method = rep(c("FDR", "MAX", "Cluster sum", "Change point"), gamma_length)
)

ggplot(df, aes(x = gamma, y = res, group = method, colour = method)) + theme_gar +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = categ.palette) +
  ylab("Bias") +
  scale_x_continuous(breaks = gamma_vec)
```

#### MAE

```{r}
df <- tibble(res = as.vector(res.mae),
             gamma = rep(gamma_vec, each = 4),
             method = rep(c("FDR", "MAX", "Cluster sum", "Change point"), gamma_length)
)

ggplot(df, aes(x = gamma, y = res, group = method, colour = method)) + theme_gar +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = categ.palette) +
  ylab("MAE") +
  scale_x_continuous(breaks = gamma_vec)
```

#### Proportion too early

```{r}
df <- tibble(res = as.vector(res.pte),
             gamma = rep(gamma_vec, each = 4),
             method = rep(c("FDR", "MAX", "Cluster sum", "Change point"), gamma_length)
)

ggplot(df, aes(x = gamma, y = res, group = method, colour = method)) + theme_gar +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = categ.palette) +
  ylab("Proportion too early") +
  scale_x_continuous(breaks = gamma_vec)
```

#### Proportion < 40 ms

```{r}
df <- tibble(res = as.vector(res.p40),
             gamma = rep(gamma_vec, each = 4),
             method = rep(c("FDR", "MAX", "Cluster sum", "Change point"), gamma_length)
)

ggplot(df, aes(x = gamma, y = res, group = method, colour = method)) + theme_gar +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = categ.palette) +
  ylab("Proportion < 40ms") +
  scale_x_continuous(breaks = gamma_vec)
```

## References

Stevens, M.H.H. (2009) A Primer of Ecology with R, 2nd Printing. Springer, New York.






